Abstract Binding Trees

### 2015-10-24 paper.tex
brief metavariable subst implementation (resolve #3)

\newcommand\Renaming[2]{
  #2\upharpoonright #1
}
\newcommand\IsRenaming[3]{
  \Renaming{\IMode{#1}}{\IMode{#2}}%
  \leadsto\OMode{#3}%
}

@@

Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified
up to $\alpha$-equivalence, the renaming judgment is functional in its input,
and so we are justified in writing $\Renaming{\varrho}{M}$ for $N$ when
$\IsRenaming{\varrho}{M}{N}$.

@@

Variable substitution in abstract binding trees is defined inductively by a
pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{E'}$:

\begin{gather*}
  \infer{
    \IsSubst{N}{x}{y}{N}
  }{
    \IsEq{x}{y}
  }\qquad
  \infer{
    \IsSubst{N}{x}{y}{y}
  }{
    \IsApart{x}{y}
  }\\[6pt]
  \infer{
    \IsSubst{N}{x}{
      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
    }
  }{
    \HypJ{
      \IsSubst{N}{x}{M_i}{M'_i}
    }{
      \Leq{i}{n}
    }
  }\\[6pt]
  \infer{
    \IsSubst{N}{x}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{
      \App{\vartheta}{E'_0,\dots,E'_n}
    }
  }{
    \HypJ{
      \IsSubst{N}{x}{E_i}{E'_i}
    }{
      \Leq{i}{n}
    }
  }
\end{gather*}
\begin{gather*}
  \infer{
    \IsSubst{N}{x}{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{y}}{M'}
    }
  }{
    \NotIn{x}{\vec{y}} &
    \IsApart{\vec{y}}{\FV{N}} &
    \IsSubst{N}{x}{M}{M'}
  }\qquad
  \infer{
    \IsSubst{N}{x}{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }
  }{
    \Member{x}{\vec{y}} &
    \IsApart{\vec{y}}{\FV{N}}
  }
\end{gather*}

@@

Metavariable substitution is defined inductively by the judgment
$\IsSubst{E}{\MV{m}}{M}{M'}$:

\begin{gather*}
  \infer{
    \IsSubst{E}{\MV{m}}{x}{x}
  }{
  }\\[6pt]
  \infer{
    \IsSubst{E}{\MV{m}}{
      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{n}}{\vec{u}}{M'_0,\dots,M'_n}
    }
  }{
    \IsApart{\MV{m}}{\MV{n}} &
    \HypJ{
      \IsSubst{E}{\MV{n}}{M_i}{M'_i}
    }{
      \Leq{i}{n}
    }
  }\\[6pt]
  %%% \infer{
  %%%   \IsSubst{
  %%%     \MkBTm{\vec{u}}{x_0,\dots,x_n}{N}
  %%%   }{\MV{m}}{
  %%%     \MApp{\MV{n}}{\vec{v}}{M_0,\dots,M_n}
  %%%   }{
  %%%     %TODO!!!
  %%%     %\Rename{\vec{v}}{\vec{u}}{
  %%%     \Subst{\vec{M}}{\vec{x}}{N}
  %%%     % }
  %%%   }
  %%% }{
  %%%   \IsEq{\MV{m}}{\MV{n}}
  %%% }\\[6pt]
  %%% \infer{
  %%%   \IsSubst{E}{\MV{m}}{
  %%%     \App{\vartheta}{F_0,\dots,F_n}
  %%%   }{
  %%%   }
  %%% }{
  %%% }
\end{gather*}

>>>>

\newcommand\Rename[2]{
  #2\upharpoonright #1
}
\newcommand\IsRenaming[3]{
  \Rename{\IMode{#1}}{\IMode{#2}}%
  \leadsto\OMode{#3}%
}

@@

Above, the notation $\varrho\setminus\vec{u}$ means the omission of the
variables $\vec{u}$ from the renaming $\varrho$. Because terms are identified
up to $\alpha$-equivalence, the renaming judgment is functional in its input,
and so we are justified in writing $\Rename{\varrho}{M}$ for $N$ when
$\IsRenaming{\varrho}{M}{N}$.

@@

Variable substitution in abstract binding trees is defined inductively by a
pair of judgments, $\IsSubst{N}{x}{M}{M'}$ and $\IsSubst{N}{x}{E}{F}$:

\begin{gather*}
  \infer{
    \IsSubst{N}{x}{y}{N}
  }{
    \IsEq{x}{y}
  }\qquad
  \infer{
    \IsSubst{N}{x}{y}{y}
  }{
    \IsApart{x}{y}
  }\\[6pt]
  \infer{
    \IsSubst{N}{x}{
      \MApp{\MV{m}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{m}}{\vec{u}}{M'_0,\dots,M'_n}
    }
  }{
    \HypJ{
      \IsSubst{N}{x}{M_i}{M'_i}
    }{
      \Leq{i}{n}
    }
  }\\[6pt]
  \infer{
    \IsSubst{N}{x}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{
      \App{\vartheta}{F_0,\dots,F_n}
    }
  }{
    \HypJ{
      \IsSubst{N}{x}{E_i}{F_i}
    }{
      \Leq{i}{n}
    }
  }
\end{gather*}
\begin{gather*}
  \infer{
    \IsSubst{N}{x}{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{y}}{M'}
    }
  }{
    \NotIn{x}{\vec{y}} &
    \IsApart{\vec{y}}{\FV{N}} &
    \IsSubst{N}{x}{M}{M'}
  }\qquad
  \infer{
    \IsSubst{N}{x}{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{y}}{M}
    }
  }{
    \Member{x}{\vec{y}} &
    \IsApart{\vec{y}}{\FV{N}}
  }
\end{gather*}

@@

Metavariables are substituted by bound terms; since a metavariable may only
appear in an application expression $\MApp{\MV{m}}{\cdots}{\cdots}$, we will
instantiate the bound term at the supplied parameters and arguments.
Substitution for metavariables is defined inductively by the judgments
$\IsSubst{E}{\MV{m}}{M}{N}$ and $\IsSubst{E}{\MV{m}}{F}{F'}$:

\begin{gather*}
  \infer{
    \IsSubst{E}{\MV{m}}{x}{x}
  }{
  }\\[6pt]
  \infer{
    \IsSubst{E}{\MV{m}}{
      \MApp{\MV{n}}{\vec{u}}{M_0,\dots,M_n}
    }{
      \MApp{\MV{n}}{\vec{u}}{N_0,\dots,N_n}
    }
  }{
    \IsApart{\MV{m}}{\MV{n}} &
    \HypJ{
      \IsSubst{E}{\MV{n}}{M_i}{N_i}
    }{
      \Leq{i}{n}
    }
  }\\[6pt]
  \infer{
    \IsSubst{
      \MkBTm{\vec{u}}{\vec{x}}{N}
    }{\MV{m}}{
      \MApp{\MV{n}}{\vec{v}}{\vec{M}}
    }{
      N''
    }
  }{
    \IsEq{\MV{m}}{\MV{n}} &
    \IsSubst{\vec{M}}{\vec{x}}{N}{N'} &
    \IsRenaming{\MkSet{\vec{u}\mapsto\vec{v}}}{N'}{N''}
  }\\[6pt]
  \infer{
    \IsSubst{E}{\MV{m}}{
      \App{\vartheta}{F_0,\dots,F_n}
    }{
      \App{\vartheta}{F'_0,\dots,F'_n}
    }
  }{
    \HypJ{
      \IsSubst{E}{\MV{m}}{F_i}{F'_i}
    }{
      \Leq{i}{n}
    }
  }\\[6pt]
  \infer{
    \IsSubst{E}{\MV{m}}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{u}}{\vec{x}}{N}
    }
  }{
    \IsApart{\vec{u}}{\FV{E}} &
    \IsSubst{E}{\MV{m}}{M}{N}
  }
\end{gather*}

As before, we will write $\Subst{E}{\MV{m}}{M}$ for $N$ when
$\IsSubst{E}{\MV{m}}{M}{N}$.
